---
title: "Citi Bike Analysis"
output: html_notebook
---
```{r}
library(ggplot2)
library(data.table)
library(dplyr)
library(stats)
```

#### Import Data
```{r}
rm(list=ls())
# Load the cleaned data set created by citi_bike_anlaysis.py
db <- read.csv("cleaned_data.csv")

head(db, n=5)
```
```{r}
#Look at summary statistics for selected variables
db_select <- db %>% select(membership_ind, total_precip, precip_ind,trip_time, day, distance, TMAX,TMIN)
sapply(db_select, summary)
```

#### Create a Logistic Regression Model 
```{r}
#Create a logistic regression model to test what variables impact if the rider is a member
set.seed(0423)
# Select sample set
sample <- sample.int(n = nrow(db), size = floor(.80*nrow(db)), replace = F)

# training data set
db_train <- db[sample, ]

# test data set
db_test  <- db[-sample, ]
```



```{r}
# Note: I initally ran this model and included the maximum temperature and trip length (in minutes) but there was a near zero coefficent on those variables so they were removed for a more accurante model.

logistic <- glm(membership_ind ~ rush_ind + distance + precip_ind, family=binomial(link='logit'), data=db_train)
# summary of estimated model
summary(logistic)
# maximized log-likelihood
logLik(logistic)
```
```{r}
## show odds ratios
exp(coef(logistic))
## odds ratios and 95% Confidence Interval
exp(cbind(OR = coef(logistic), confint(logistic)))
```
This model suggests that distance, if the trip is during rush hour, and precipitation are statistically significant factors that predict if the rider has a membership.The odds ratio on the distance suggests that as the trip distance decreases by one kilometer, the rider is .94 less likely to be a member. The rush indicator suggests that if the trip was taken during rush hour,the the rider is 1.21 times as likely to be a member The precipitation indicator suggests that if the day has precipitation, the rider is 1.12 times as likely to have a membership.

```{r}
# Predict
db_test$predict_probability_temp <- predict(logistic, newdata = db_test, type = "response") 
db_test$predict_probability = as.matrix(db_test$predict_probability_temp)
# Actual proportion of customers who accepted the offer
cat('Actual mean:', mean(db_test$membership_ind))

```
```{r}
cat('Predicted mean:',mean(db_test$predict_probability,,na.rm = TRUE))
```

The actual mean and predicted mean were very close, indicating that the rush hour, precipitation,and distance are factors that influence if the rider has a citi bike membership. The closeness of the predicted mean to the actual mean shows that the model did a good job at predicting membership based on the variables. This information could be valuable for  operations and marketing for citi bike and could be used to try to increase rider membership.

